---
title: "R Notebook"
output: html_notebook
---
# Anomaly Detection 


```{r}
install.packages("backports")
```

```{r}
# Installing anomalize package
# ---
# 
#install.packages("anomalize")
```

```{r}
#2. Reading the Data

# Install the following packages:
install.packages("foreign")
library(foreign)
install.packages("car")
install.packages("Hmisc")
install.packages("reshape")
```

```{r}
install.packages("lubridate")
install.packages("tidyr")
library(tidyr)
library(lubridate)
```



```{r}
#install.packages("anomalize")
devtools::install_github("tidyverse/tidyverse")
```

```{r}
 devtools::install_github("business-science/anomalize")
```


```{r}
library(anomalize) #tidy anomaly detectiom
library(tidyverse) #tidyverse packages like dplyr, ggplot, tidyr
library(coindeskr) #bitcoin price extraction from coindesk
```


```{r}
neededPackages = c("devtools", "Rcpp", "ggplot2")
for (i in 1:length(neededPackages)) {
    ifelse(!neededPackages[i] %in% installed.packages(), install.packages(neededPackages[i]), 
    print(paste(neededPackages[i], "is already installed.")))
}
```


```{r}
library(devtools)
```

```{r}
library(AnomalyDetection)
```


Prepare the data for the AnomalyDetection package by keeping only the date and page views and discard all other variables. The input must be a series of pairs. Check the structure of the higginsWikipedia dataframe. The date column must be in POSIXct format.
```{r}
columns_to_keep=c("Date","Sales")
Sales_Forecasting=Sales_Forecasting[,columns_to_keep]
str(Sales_Forecasting)
```
AnomalyDetection Method
Apply anomaly detection and plot the results.
```{r}
AnomalyDetectionSales_Forecasting = AnomalyDetectionTs(Sales_Forecasting, direction="pos", plot=TRUE, title = "Anomaly Detection")
AnomalyDetectionSales_Forecasting$plot
```

Look at the dates the anomalies occured.
```{r}
#Look at the dates the anomalies occured.
AnomalyDetectionHiggins$anoms$timestamp
```
```{r}
#devtools::install_github("business-science/anomalize")
```


```{r}
#install.packages('anomalize')
#Update from github
#library(devtools)
#install_github("business-science/anomalize")
#Load the package
library(anomalize)
library(tidyverse)
```
```{r}
rstudioapi::getThemeInfo()
```

```{r}
# install.packages("remotes")
remotes::install_github("r-lib/rlang")
```

```{r}
#install.packages("rlang")
```

```{r}
sessionInfo()
```
```{r}
#install.packages("tidyquant")
#install.packages("timetk")
#install.packages("knitr")
```


```{r}
#install.packages("timetk")
```


```{r}
library(tidyquant)  # Used for business-ready ggplot themes
library(anomalize)  # Identify and clean time series anomalies
library(knitr)      # For kable() function
```


```{r}
rlang::last_error()
```

```{r}
install.packages("tsibble")
```

```{r}
library(tsibble)
```


```{r}

# Grouping by server and converting to tibbletime
Sales_Forecasting <- read_csv("Supermarket_Sales_Forecasting.csv") %>%
  group_by(Sales)
 

Sales_Forecasting
```

```{r}
Sales_Forecasting$Date<- as.Date(Sales_Forecasting$Date)
str(Sales_Forecasting$Date)
```


```{r}
library(dplyr)
library(tibbletime)
```
```{r}
prep_tbl_time(Sales_Forecasting)
```


```{r}
data_tbl <- tibble(
    date  = seq.Date(from = as.Date("0001-05-20"), by = "day", length.out = 10),
    value = rnorm(10)
    )

prep_tbl_time(data_tbl)
#as_tbl_time(logs)
#prep_tbl_time(logs)
```

```{r}
Sales_Forecasting %>%
    ggplot(aes(Date, Sales)) +
    geom_point(color = "#2c3e50", alpha = 0.25) +
    facet_wrap(~ Sales, scale = "free_y", ncol = 1) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    labs(title = "Sales data plot",
         subtitle = "Grouping by Sales")
```

```{r}
Sales_Forecasting %>%
    time_decompose(Sales) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 2, alpha_dots = 0.5)
```

```{r}
Sales_Forecasting<-as.data.frame(Sales_Forecasting$Sales)
Sales_Forecasting
```
Plot the data using ggplot2. Set the x-axis to the Date and the y-axis to the number of views.
```{r}
library(ggplot2)
ggplot(Sales_Forecasting, aes(x=Date, y=Sales, color=Sales)) + 
  geom_line()
```


